A unified model for parsing and sentence understanding

### 2016-03-14 writing/hard_stack_paper/paper.tex
Add shift-reduce parsing intro

\section{The transition-based sentence model}

\subsection{Intuition: Transition-based parsing}

The PINN is based on a formalism adapted from transition-based parsing (\question{What's a good classic citation for this?}).

\todo{[JG, SB] Brief into to transition-based parsing.}

@@

The full PINN model is shown in Figure~\ref{fig:model:1d}. The PINN takes the form of a transition-based parser, but is designed to produce sentence representations given parse structures, rather than producing parse structures as its output. It extends the standard transition-based model in the following ways:
\begin{itemize}
\item The intermediate representations in the parser, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\item After processing a sentence, the model has a single vector representation of the sentence on its stack. This representation is a differentiable function of all prior representations. We can thus learn the weights of the model via backpropagation of some downstream error.
\end{itemize}

>>>>

\section{The shift-reduce sentence model}

\subsection{Shift-reduce parsing}

The PINN is inspired by the shift-reduce parsing formalism \cite{aho1986compilers}, which builds a tree structure over a sequence (e.g. a natural language sentence) by a single left-to-right scan over the tokens of the sequence. Shift-reduce parsing has been successfully applied in algorithms for natural language parsing \cite{nivre2003efficient}.

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. (For a binary-branching tree structure, this requires $T = 2N - 1$ operations.)

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below $\mid$ denotes an infix \textit{cons} function.
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

The core PINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output (as opposed to a tree). This is accomplished by the following modifications to the shift-reduce algorithm:
\begin{itemize}
\item The intermediate representations in the parser, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\end{itemize}
